---
output:
  pdf_document: default
  html_document: default
---

# Regression Models Course Project - Motor Trend Project

RB,   
08.02.2021

```{r results='hide', message=FALSE, warning=FALSE}
library(datasets)
library(ggplot2)
```

# Introduction
In this analysis, exploring the mtcars data set we will answer the following questions using amongst other things linear regression and multivariable linear regression:

**1. Is an automatic or manual transmission better for miles per gallone (MPG)?**

**2. What is the MPG difference between automatic and manual transmissions?**  


## Exploratory Analysis
```{r}
data(mtcars)
head(mtcars)
```

Transform variables am, cyl, vs, am, gear, carb, into factor variable with discrete values:

```{r}
mtcars$cyl <- factor(mtcars$cyl)
mtcars$vs <- factor(mtcars$vs)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
mtcars$am <- factor(mtcars$am,labels=c('Automatic','Manual'))
str(mtcars)
```

We calculate and plot (Fig. 1) the averaged miles per gallon for manual and automatic transmission each
```{r}
aggregate(mpg~am, data = mtcars, mean)
```

```{r}
boxplot(mpg ~ am, data = mtcars, col = (c("green","blue")), ylab = "Miles Per Gallon (MPG)", xlab = "Type of Transmission")
```

## Statistical Testing 
Based on our explorative data analysis we will now test if the calculated MPG difference of **7.15** between manual and automatic is significant. 
To test our hypothesis we use a two-sided-unpaired Welch t-test assuming unequal variance with significance level 0.05.

```{r}
automatic <- mtcars[mtcars$am == "Automatic",]
manual <- mtcars[mtcars$am == "Manual",]
t.test(automatic$mpg, manual$mpg, conf.level = 0.95, 
       alternative="two.sided", var.equal = FALSE, paired = FALSE)
```

A p-value of **0.0013** was obtained and the confidence interval of the test does not contain zero. We conclude that we reject the null hypothesis. 
This means that there is significant difference between the
the transmission types manual and automatic.


## Regression Analysis - Linear Models

First we start with a simple linear regression model by taking "am" as the only regressor to predict mpg:
```{r}
model_base <- lm(mpg ~ am, data = mtcars)
summary(model_base)
```

A adjusted R-squared value of is obtained which means that this simple linear model explains only **34%** of the variability.

So the next step is to use adjusted multivariable linear regression model including **cyl**, **disp**, **hp** and **wt** as counfounding variables.


```{r}
model_adjust <- lm(mpg ~ am + cyl + disp + hp + wt, data = mtcars)
summary(model_adjust)
```
 
Now a adjusted R-squared value of **83%** is obtained which is much higher than for the simple model. Lets see what the ANOVA-analysis shows.


```{r}
anova(model_base, model_adjust)
```
We obtain a very significant p-value of **8.6e-8**, hence we conclude that the adjusted model predicts the mpg significantly better than the simple model.

## Residuals and Diagnostics

```{r}
par(mfrow = c(2, 2))
plot(model_adjust)
```

Residuals vs. Fitted plot show that the points are randomly distributed on the plot verifying the independence condition.
In the normal Q-Q plot the points fall mostly on the line indicating that the residuals are normally distributed.

## Conclusion

The analyis indicates that cars with manual transmission achieve a higher MPG (**+ 7.15**) compared to automatic cars (Fig. 1). 
However when confounding variables like number of cylinders, displacement, gross horsepower and weight are taken into account the difference between manual and automatic transmission (**1.8**, adjusted by cyl, disp, hp, and wt) is not as clear as the explorative analyis was suggesting in the beginning. 
